---
title: "Resource competition"
author: "ZZ"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    number_sections: true
    toc_float:
      collapsed: no
      smooth_scroll: no
---
\


**The question**:\
Here, we aim to test whether **resource availability** (light and/or nutrient) affects the competitive outcomes between **alien** and **native** plants.

**The design**:\
In Konstanz, Germany, Angelina grew 16 species (8 aliens and 8 natives), either alone, or with intraspecific competition, or with interspecific competition (32 out of the 64 alien-native pairs were included). \

In Changchun, China, Yanjie's group did a similar experiment with 8 species (4 aliens and 4 natives). They also included alien-alien and native-native pairs.\


**The analyses**:\
First, I pooled two experiments together, with the alien-alien and native-native pairs in Yanjie's experiment removed (model: m0).\
Then, I analyzed the two experiments seperately (models: m_zz & m_yj).\

**The results**:\
Most importantly, all three models showed that aliens were more competitive than natives, when both nutrient and light availabilities were high. This is indicated by the significant, or margnially significant, interactions between status (alien or native), fertilization and shading.


**To do**:\
1. The trade-off between light competition and nutrient competition.


<style type="text/css">

body{ /* Normal  */
      font-size: 16px;
      font-family: "Times New Roman"
  }
td {  /* Table  */
  font-size: 14px;
}
h1.title {
  font-size: 34px;
  color: Black;
  font-family: "Arial";
}
h1 { /* Header 1 */
  font-size: 26px;
  color: DarkBlue;
  font-family: "Arial";
}
h2 { /* Header 2 */
    font-size: 22px;
    font-family: "Arial";
    color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Arial";
  color: DarkBlue;
}

h4 { /* Header 4 */
  font-size: 16px;
  font-family: "Arial";
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 12px;
}
</style>


```{r}
knitr::opts_chunk$set(message = F,warning = F)
dir.create('fig', showWarnings = F)
dir.create('table', showWarnings = F)
```


```{r}
library(tidyverse)
library(Taxonstand)
library(nlme)
library(knitr)
library(kableExtra)
library(gridExtra)
library(effects)
load('results.RData')
source('functions.R')
```

# read data
```{r}
dat_yj_raw <- read.csv('Data_ABcomp_HFJ.csv')
dat_zz <- read.csv('Konstanz20210318.csv') %>% 
  # dplyr::select(- Box, -Box.2)
  rename(cage_rand1 = Box,
         cage_rand2 = Box.2)
colnames(dat_zz) <- colnames(dat_zz) %>% tolower()

dat_yj <- dat_yj_raw %>% 
  dplyr::select(Nr., Target_Species, Target_Status, Competition, Competitor_Species, Competitor_Status,
         N_Treatment, Shading_Treatment,
         Target_Shoot, Competitor_Shoot, Target_Root)

```


```{r}
# make pots with two data points into two rows
dat_yj_t <- dat_yj %>% # target individuals
  dplyr::select(-Competitor_Shoot) %>% 
  rename(biomass = Target_Shoot,
         target  = Target_Species,
         status  = Target_Status,
         comp        = Competitor_Species,
         status_comp = Competitor_Status,
         comp_treat  = Competition ,
         fert  = N_Treatment,
         light =  Shading_Treatment,
         id = Nr., 
         root = Target_Root)

dat_yj_c <- dat_yj %>% # competitors
  dplyr::select(-Target_Shoot) %>% 
  rename(biomass = Competitor_Shoot,
         target  = Competitor_Species,
         status  = Competitor_Status,
         comp        = Target_Species,
         status_comp = Target_Status,
         comp_treat  = Competition ,
         fert  = N_Treatment,
         light =  Shading_Treatment,
         id = Nr.,
         root = Target_Root) %>% 
  mutate(root = NA)


dat_yj <- rbind(dat_yj_t, dat_yj_c) %>% 
  filter(status != 'Without') %>% 
  mutate(status = ifelse(status == 'Self', as.character(status_comp), as.character(status))#,
         #status_comp = ifelse(status_comp == 'Self', as.character(status), as.character(status_comp)),
         ) %>% 
  #----- change names ---------#
  mutate(status      = fct_recode(status, 'alien' = 'Invasive', 'native' = 'Native'),
         status_comp = fct_recode(status_comp, 'alien' = 'Invasive', 'native' = 'Native',
                                  'intra' = 'Self', 'alone' = 'Without'),
         light      = fct_recode(light, 'Ambient' = 'Control'),
         comp_treat = tolower(comp_treat),
         comp = ifelse(comp == 'No_competitor', 'Non', as.character(comp))
         ) %>% 
  #----- change reference levels -------#
  mutate(light = fct_relevel(light, 'Shading'),
         fert = fct_relevel(fert, 'Low')) %>% 
  droplevels()

# remove pots contained dead (NA) plants
list <- dat_yj$id[is.na(dat_yj$biomass)] %>% unique
dat_yj <- dat_yj %>% 
  filter(!id %in% list)
```


```{r}
# standardize names
sp_list <- dat_yj$target %>% unique %>% str_replace('_', ' ')



sp_list <- TPL(sp_list) %>% 
  dplyr::select(Taxon, Family) %>% 
  mutate(Taxon = str_replace(Taxon, ' ', '_')) 

dat_yj <- dat_yj %>% 
  left_join(sp_list, by = c('target' = 'Taxon')) %>% 
  rename(family_tar = Family) %>% 
  left_join(sp_list, by = c('comp' = 'Taxon')) %>% 
  rename(family_comp = Family) %>% 
  mutate(family_comp = ifelse(is.na(family_comp), 'Non', as.character(family_comp)))

rm(sp_list, list, dat_yj_c, dat_yj_t)
```


```{r}
dat_zz <- dat_zz %>% 
  mutate(status = fct_recode(status, 'alien' = 'ex', 'native' = 'nat'),
         light  = fct_recode(light, 'Ambient' = 'W', 'Shading' = 'G'),
         fert   = fct_recode(fert , 'High' = 'H', 'Low' = 'L'),
         target = str_remove(target, '\\s\\(\\w*\\)'), # remove braces and characters between them
         comp   = str_remove(comp, '\\s\\(\\w*\\)'),
         target = str_replace(target, '\\s', '_'),
         comp   = str_replace(comp, '\\s', '_')
         ) %>% 
  #----- change reference levels -------#
  mutate(light = fct_relevel(light, 'Shading'),
         fert = fct_relevel(fert, 'Low')) %>% 
  droplevels()
```







# analyses I
## two datasets pooled
merge the two datasets (only alone, intraspecific competition, and alien-native pairs)
```{r}
dat <- dat_yj %>% 
  filter(as.character(status) != as.character(status_comp)) %>% # exclude native-native, and alien-alien combs
  dplyr::select(-status_comp) %>% 
  bind_rows(dat_zz) %>% 
  mutate(exp = str_extract(id, '^.')) %>% # distinguish the two experiments
  mutate(cage_rand1 = ifelse(is.na(cage_rand1), as.character(light), as.character(cage_rand1)),
         cage_rand2 = ifelse(is.na(cage_rand2), as.character(light), as.character(cage_rand2))
         )


```

add contrasts\
T1: alone vs competition\
T2: intra- vs inter- competition
```{r}
mat <- matrix(c(1/3, 1/3, 1/3,  
                -1,  0.5, 0.5,
                0,     1, -1), ncol=3)

mat_solve = solve(t(mat))[,2:3] %>% 
  as.data.frame() %>% 
  mutate(comp_treat = c('alone', 'inter', 'intra')) %>% 
  rename(T_alone_comp = V1,
         T_intra_inter = V2)

dat <- dat %>% 
  left_join(mat_solve)
rm(mat)

# for random slope
dat <- dat %>% 
  mutate(light_slope = paste(target, light, sep = '_'),
         fert_slope = paste(target, fert, sep = '_'))
```


```{r, eval = F}
u <- rep(1, nrow(dat))
m0 <- lme(log(biomass) ~ status*(T_alone_comp + T_intra_inter) * fert * light, # + length_in_mm
          random=list(u = pdBlocked(list(
                                        pdIdent(form = ~ exp-1),
                                        pdIdent(form = ~ family_tar-1),
                                        pdIdent(form = ~ target-1),
                                        pdIdent(form = ~ light_slope -1),
                                        pdIdent(form = ~ fert_slope -1),  # similar to random slope
                                        pdIdent(form = ~ family_comp-1),
                                        pdIdent(form = ~ comp-1),
                                        pdIdent(form = ~ cage_rand1-1)
                                        ))),
          weights = varComb(varIdent(form= ~1|target),
                            varIdent(form= ~1|comp_treat)),
          control = lmeControl(msMaxIter = 10000, msMaxEval = 10000),
          data=dat)


```


```{r}
m0 %>% Ftable
```



## seperately
### Germany (Konstanz, ZZ)
```{r}
dat_zz_only <- dat %>% 
  filter(exp == 'A')
```



```{r, eval = F}
u <- rep(1, nrow(dat_zz_only))
m_zz <- lme(log(biomass) ~ status*(T_alone_comp + T_intra_inter) * fert * light, 
            random=list(u = pdBlocked(list(
                                          pdIdent(form = ~ family_tar-1),
                                          pdIdent(form = ~ target-1),
                                          pdIdent(form = ~ family_comp-1),
                                          pdIdent(form = ~ light_slope-1),
                                          pdIdent(form = ~ fert_slope-1),
                                          pdIdent(form = ~ comp-1)
                                          ))), # adding pot id or block does not improve the model, but consume lots of computation.
            weights = varComb(varIdent(form= ~1|target),
                              varIdent(form= ~1|comp_treat)
                              ),
            control = lmeControl(msMaxIter = 100000),
            data=dat_zz_only)

```


```{r}
m_zz %>% Ftable
```



### China (Changchun, YJ)
As there is competition between aliens and between natives, another contrast - alien vs native competitor - was added. 
```{r}
mat <- matrix(c(1/4, 1/4, 1/4, 1/4,  
                -1,  1/3, 1/3, 1/3, # alone vs comp
                0,   1,   -0.5, -0.5,  # 
                0,   0,   1,  -1), # native vs alien
              ncol = 4)

mat_solve_yj = solve(t(mat))[,2:4] %>% 
  as.data.frame() %>% 
  mutate(status_comp = c('alone', 'intra', 'alien', 'native')) %>% 
  rename(T_alone_comp   = V1,
         T_intra_inter  = V2,
         T_alien_native = V3
         )

dat_yj_only <- dat_yj %>% 
  left_join(mat_solve_yj) %>% 
  mutate(light_slope = paste(target, light, sep = '_'),
         fert_slope = paste(target, fert, sep = '_'))
rm(mat)
```


```{r, eval = F}
u <- rep(1, nrow(dat_yj_only))
m_yj <- lme(log(biomass) ~ status*(T_alone_comp + T_intra_inter + T_alien_native) * fert * light, # + length_in_mm
            random=list(u = pdBlocked(list(
                                          pdIdent(form = ~ family_tar-1),
                                          pdIdent(form = ~ target-1),
                                          pdIdent(form = ~ family_comp-1),
                                          pdIdent(form = ~ light_slope-1),
                                          pdIdent(form = ~ fert_slope-1),
                                          pdIdent(form = ~ comp-1)
                                          ))),
          weights = varComb(varIdent(form= ~1|target),
                            varIdent(form= ~1|status_comp) # removing this is ok
                            ),
          control = lmeControl(msMaxIter = 10000),
          data=dat_yj_only)

```



```{r}
m_yj %>% Ftable
```


Export all tables in one file

```{r}
tables <- lapply(list(m_yj, m_zz, m0), Ftable_ext)
table_in_one <- tables[[1]] %>% # China
  left_join(tables[[2]], by = 'items') %>% # Germany 
  full_join(tables[[3]], by = 'items') %>%  # both
  replace_na(list(`.y` = ' ',
                  ` ` = ' '
                  )) %>% 
  mutate_if(is.numeric, round,3)
  

# table_in_one %>% kable() %>% 
#   kable_styling()
```


```{r}
write.csv(table_in_one, 'table/table_in_one.csv', fileEncoding = "UTF-8")
```




# Figures for analysis I
All figures are ploted with raw data.\

## status:fertilization:shading
Aliens are more competitive than native when both nutrient and light levels are high.\


In Yanjie's experiment, the boxplot based on raw data shows that aliens were less competitive than natives, when nutrient level is low and light level is high. This is not consistent with the other two figures. But if we use modelled values, the patterns will be the same.\

So, I would prefer to plot with modelled values.
```{r}
theme <-theme(panel.background = element_blank(),
              panel.border     = element_rect(fill = NA), 
              strip.background = element_blank(),
              strip.text.x = element_text(size = 10),
              axis.text.x = element_text(size = 8, angle = 0, hjust = 0.5, colour = 'black'),
              axis.text.y  = element_text(size = 8, colour = 'black'),
              axis.title = element_text(size = 10),# size of x  y lab
              #legend.position   = c(0.9, 0.85),
              legend.title      = element_text(size = 8),
              legend.text       = element_text(size = 6),
              legend.key.height = unit(0.8,'line'),
              legend.background = element_blank(),
              legend.key        = element_rect(colour = NA, fill = NA),
              plot.tag    = element_text(size = 8, face  = 'bold'),
              plot.title  = element_text(hjust = 0.5, size = 12),
              plot.margin = margin(t = 0.5, r = 0.1, b = 0, l = 0.1, unit = "cm"))
```



modelled values
```{r}
u <- rep(1, nrow(dat))
m_fig   <- update(m0, .~ status * fert * light)
m_light <- update(m0, ~ light * comp_treat)
m_fert  <- update(m0, ~ fert * comp_treat)
```


```{r}
u <- rep(1, nrow(dat_yj))
m_status <- update(m_yj, .~ status * status_comp)
```



## origin:fert:light
```{r}
t <- effect('status:fert:light', m_fig) %>% as.data.frame() %>% 
  mutate(xaxis = c(1, 2, 4, 5, 7, 8, 10, 11))

dat_fig1 <- dat %>% 
  left_join(t %>% dplyr::select(status, light, fert, xaxis))
```



```{r}
p1 <- dat_fig1 %>% 
  mutate(origin = status) %>% 
  ggplot() +
  geom_point(aes(x = xaxis, y = log(biomass)), 
             position = position_jitter(w = 0.15, h = 0), 
             col = 'black', 
             alpha = 0.04, 
             size = 0.6) +
  geom_point(data = t,
             aes(x = xaxis, y = (fit), col = status), 
             size = 1.5) +
  geom_errorbar(data = t, 
                aes(ymin = (fit-se), ymax = (fit+se), x = xaxis, col = status), 
                width = 0,  inherit.aes = FALSE,
                size = 0.8) +
  scale_x_continuous(limits = c(0.5, 11.5), breaks = c(1.5, 4.5, 7.5, 10.5), 
                     labels = c("Low resources", "+ nutrient", '+ light', '+ nutrient & + light')) + 
  scale_y_continuous(limits = c(-4, 4)) + 
  # scale_color_manual(values = col_fig) + 
  labs(x = '', y = 'ln(Aboveground biomass)', col = 'Origin')+ theme
p1

```


```{r}
ggsave('./fig/pooled.pdf', p1, height = 7/2.54, width = 11/2.54, useDingbats=FALSE)
```



## fert:competition
```{r}
t_fert <- effect('fert:comp_treat', m_fert) %>% as.data.frame() %>% 
  arrange(desc(fert)) %>% 
  mutate(xaxis = c(1:3, 5:7))

dat_fig2a <- dat %>% 
  left_join(t_fert %>% dplyr::select(fert, comp_treat, xaxis))
```

```{r}
col_fig <- c('#F39C12', '#115888', '#85C1E9')

p2a <- dat_fig2a %>% 
  ggplot() +
  geom_point(aes(x = xaxis, y = log(biomass)), 
             position = position_jitter(w = 0.15, h = 0), 
             col = 'black', 
             alpha = 0.04, 
             size = 0.6) +
  geom_point(data = t_fert, 
             aes(x = xaxis, y = (fit), col = comp_treat), 
             size = 1.5) +
  geom_errorbar(data = t_fert, 
                aes(ymin = (fit-se), ymax = (fit+se), x = xaxis, col = comp_treat), 
                width = 0,  inherit.aes = FALSE,
                size = 0.8) +
  scale_x_continuous(limits = c(0.5, 7.5), breaks = c(2, 6), 
                     labels = c("Low nutrient", "+ nutrient")) + 
  scale_y_continuous(limits = c(-4, 4)) + 
  # scale_y_continuous(limits = c(0, 20)) + 
  scale_color_manual(values = col_fig) + 
  labs(x = '', y = 'ln(Aboveground biomass)', tag = 'a',  col = 'Competition')+ theme
p2a
```


## light:competition
```{r}
t_light <- effect('light:comp_treat', m_light) %>% as.data.frame() %>% 
  arrange(desc(light)) %>% 
  mutate(xaxis = c(1:3, 5:7))

dat_fig2b <- dat %>% 
  left_join(t_light %>% dplyr::select(light, comp_treat, xaxis))
```

```{r}
p2b <- dat_fig2b %>% 
  ggplot() +
  geom_point(aes(x = xaxis, y = log(biomass)), 
             position = position_jitter(w = 0.15, h = 0), 
             col = 'black', 
             alpha = 0.04, 
             size = 0.6) +
  geom_point(data = t_light, 
             aes(x = xaxis, y = (fit), col = comp_treat), 
             size = 1.5) +
  geom_errorbar(data = t_light, 
                aes(ymin = (fit-se), ymax = (fit+se), x = xaxis, col = comp_treat), 
                width = 0,  inherit.aes = FALSE,
                size = 0.8) +
  scale_x_continuous(limits = c(0.5, 7.5), breaks = c(2, 6), 
                     labels = c('Low light', "+ light")) + 
  scale_y_continuous(limits = c(-4, 4)) + 
  # scale_y_continuous(limits = c(0, 20)) + 
  scale_color_manual(values = col_fig) + 
  labs(x = '', y = 'ln(Aboveground biomass)', tag = 'b', col = 'Competition') + 
  theme +
  theme(
        legend.text = element_text(color = "white"),
        legend.title = element_text(color = "white"),
        legend.key = element_rect(fill = "white")
    )
p2b
```


```{r}

ggsave('./fig/comp.pdf', height = 7/2.54, width = 18/2.54,
       grid.arrange(p2a, p2b, ncol = 2))
```


## origin:origin_comp
```{r}
t_status <- effect('status:status_comp', m_status) %>% as.data.frame() %>% 
  arrange(status)
t_status <- t_status[c(1,4, 5, 8),] %>% 
  mutate(xaxis = c(1:2, 4:5))

dat_fig3 <- dat_yj_only %>% 
  left_join(t_status %>% dplyr::select(status, status_comp, xaxis)) %>% 
  filter(!is.na(xaxis))
```

```{r}
p3 <- dat_fig3 %>% 
  ggplot() +
  geom_point(aes(x = xaxis, y = log(biomass)), 
             position = position_jitter(w = 0.15, h = 0), 
             col = 'black', 
             alpha = 0.04, 
             size = 0.6) +
  geom_point(data = t_status, 
             aes(x = xaxis, y = (fit), col = status), 
             size = 1.5) +
  geom_errorbar(data = t_status, 
                aes(ymin = (fit-se), ymax = (fit+se), x = xaxis, col = status), 
                width = 0,  inherit.aes = FALSE,
                size = 0.8) +
  scale_x_continuous(limits = c(0.5, 5.5), breaks = c(1, 2, 4, 5), 
                     labels = c('Alien', 'Native', 'Alien', 'Native')) + 
  scale_y_continuous(limits = c(-4, 4)) + 
  # scale_y_continuous(limits = c(0, 20)) + 
  labs(x = '\nOrigin of the competitor', y = 'ln(Aboveground biomass)', col = 'Origin of the target')+ theme
p3
```

```{r}
ggsave('./fig/origin.pdf', p3, height = 7/2.54, width = 11/2.54, useDingbats=FALSE)
```

## old figures
```{r, fig.width = 5, fig.height = 3.5}
col_fig <- c('#f7ad4d', '#b041b0')
dat %>% 
  filter(biomass > 0.001) %>% 
  ggplot(aes(x = fert, y = log(biomass), fill = status)) + geom_boxplot() +
  scale_fill_manual(values = col_fig) + 
  facet_wrap( ~ light) + labs(title = 'Pooled', y = '(Aboveground biomass)\n', x = '\nNutrient level') + theme



dat_zz_only %>% 
  ggplot(aes(x = fert, y = log(biomass), fill = status)) + geom_boxplot() +
  scale_fill_manual(values = col_fig) + 
  facet_wrap( ~ light) + labs(title = 'Germany', y = 'Aboveground biomass\n', x = '\nNutrient level') + theme


dat_yj_only %>% 
  ggplot(aes(x = fert, y = log(biomass), fill = status)) + geom_boxplot() +
  scale_fill_manual(values = col_fig) + 
  facet_wrap( ~ light) + labs(title = 'China', y = 'Aboveground biomass\n', x = '\nNutrient level') + theme
```


```{r, fig.width = 5, fig.height = 3.5}
dat %>% 
  ggplot(aes(x = comp_treat, y = log(biomass))) + geom_boxplot() +
  facet_wrap( ~ light) + labs(title = 'Pooled', y = 'Aboveground biomass\n', x = '\nCompetition') + theme

```


```{r, fig.width = 5, fig.height = 3.5}
dat %>% 
  mutate(fert = ifelse(fert == 'Low', 'Low nutrient', 'High nutrient')) %>% 
  ggplot(aes(x = comp_treat, y = log(biomass))) + geom_boxplot() +
  facet_wrap( ~ fert) + labs(title = 'Pooled', y = 'Aboveground biomass\n', x = '\nCompetition') + theme
```


```{r, fig.width = 5, fig.height = 3.5}
dat %>% 
  ggplot(aes(x = fert, y = biomass)) + geom_boxplot() +
  facet_wrap( ~ light) + labs(title = 'Pooled', y = 'Aboveground biomass\n', x = '\nNutrient level') + theme
```

only significant in Yanjie's experiment. \
Aliens were less limited by natives than by aliens, but this does not hold for natives.
```{r, fig.width = 5, fig.height = 3.5}
dat_yj_only %>% 
  filter(status_comp %in% c('alien', 'native')) %>% 
  ggplot(aes(x = status, y = log(biomass), fill = status_comp)) + geom_boxplot() + theme + labs(title = 'China')
```

## status:intra_inter:light
only significant in the Konstanz-experiment.\
Under high-level-light condition, aliens had higher biomass than natives when growing with each other. This is not true for intraspecific competition and not true under low-level-light condition. 
```{r, fig.width = 5, fig.height = 3.5}
dat_zz_only %>% 
  filter(comp_treat != 'alone') %>% 
  ggplot(aes(x = comp_treat, y = biomass, fill = status)) + geom_boxplot() + 
  facet_wrap(~light) + theme + labs(title = 'Germany')
```



export tables
```{r}

m0 %>% Ftable %>% save_kable('table/pooled.png')
m_zz %>% Ftable %>% save_kable('table/zz_single.png')
m_yj %>% Ftable %>% save_kable('table/yj_single.png')

```



# analysis II
proportion of alien
```{r}
dat_inter <- dat %>% 
  filter(comp_treat == 'inter')
dat_ratio <- dat_inter %>% 
  filter(status == 'alien') %>% 
  left_join(dat_inter %>% filter(status == 'native') %>% select(id, biomass) %>% rename(bio_native = biomass)) %>% 
  mutate(ratio = biomass/(bio_native),
         percent = biomass/(biomass + bio_native))

rm(dat_inter)

```

```{r}
u <- rep(1, nrow(dat_ratio))
m_ratio <- lme(log(ratio) ~ fert * light, 
            random=list(u = pdBlocked(list(
                                          pdIdent(form = ~ family_tar-1),
                                          pdIdent(form = ~ target-1),
                                          pdIdent(form = ~ family_comp-1),
                                          pdIdent(form = ~ light_slope-1),
                                          pdIdent(form = ~ fert_slope-1),
                                          pdIdent(form = ~ comp-1)
                                          ))), # adding pot id does not improve the model, but consume lots of computation.
            weights = varComb(varIdent(form= ~1|target),
                              varIdent(form= ~1|comp_treat)
                              ),
            control = lmeControl(msMaxIter = 100000),
            data=dat_ratio)
```


```{r}
m_ratio %>% Ftable
```




```{r}
dat_ratio %>% 
  ggplot(aes(x = fert, y = log(ratio))) + geom_boxplot() +
  facet_wrap( ~ light) + labs(title = 'Pooled', y = 'alien/native\n', x = '\nNutrient level') + theme

```





# analysis III
root allocation
```{r}
dat_root <- dat %>% filter(!is.na(root)) %>% 
  mutate(RSR = log(root/biomass))
```

```{r}
u <- rep(1, nrow(dat_root))
m_root <- lme(RSR ~ status*fert * light, 
            random=list(u = pdBlocked(list(
                                          pdIdent(form = ~ family_tar-1),
                                          pdIdent(form = ~ target-1),
                                          pdIdent(form = ~ light_slope-1),
                                          pdIdent(form = ~ fert_slope-1)
                                          ))), # adding pot id does not improve the model, but consume lots of computation.
            weights = varComb(varIdent(form= ~1|target)
                              ),
            control = lmeControl(msMaxIter = 100000),
            data=dat_root)
```



```{r}
m_root %>% Ftable()
```

```{r}
dat_root %>% 
  ggplot(aes(x = fert, y = RSR, fill = status)) + geom_boxplot() +
  facet_wrap( ~ light) + labs(title = 'China', y = 'RSR\n', x = '\nNutrient level') + theme
```






# analysis IIII

Increasing one resource is always testing how much it limits the growth. But there are two circumstances: whether the other one is limited.


?min requirement, or max growth
```{r}
# calculate the mean
test <- dat %>% 
  filter(comp_treat == 'alone') %>% 
  group_by(target, light, fert, exp, family_tar, status) %>% 
  summarise(mean = mean(log(biomass)))


test <- test %>% 
  #--------- add shading-low as the reference level------#
  left_join(test %>% 
              filter(light == 'Shading', fert == 'Low') %>% # low as ck
              rename(mean_ck_low = mean) %>% 
              ungroup %>% 
              dplyr::select(target, mean_ck_low)) %>% 
  #-----------add ambient-high as the reference level--------------------#
    left_join(test %>% 
              filter(light == 'Ambient', fert == 'High') %>%  # high as ck
              rename(mean_ck_high = mean) %>% 
              ungroup %>% 
              dplyr::select(target, mean_ck_high)) %>% 
  #-------------------------------------#
  mutate(lnRR_low = mean - mean_ck_low,
         lnRR_high = mean_ck_high - mean)

test2 <- test %>% 
  filter((fert == 'Low' & light == 'Ambient')|(fert == 'High' & light == 'Shading')) %>% 
  mutate(treat = paste(fert, light, sep = '_')) %>% 
  ungroup %>% 
  dplyr::select(-light, -fert, -mean) %>% 
  pivot_wider(names_from = treat, values_from = c(lnRR_low, lnRR_high)) %>% 
  rename(nutrient_min = `lnRR_low_High_Shading`,
         light_min    = `lnRR_low_Low_Ambient`,
         light_max    = `lnRR_high_High_Shading`,
         nutrient_max = `lnRR_high_Low_Ambient`
         )

# test3 <- test %>% 
#   filter((fert == 'Low' & light == 'Ambient')|(fert == 'High' & light == 'Shading')) %>% 
#   mutate(treat = paste(fert, light, sep = '_')) %>% 
#   ungroup %>% 
#   pivot_longer(cols = starts_with('ln'),
#                names_to = 'type',
#                values_to = 'lnRR') %>% 
#   mutate()

```

```{r}
test2 %>% ggplot() +
  aes(x = nutrient_min, y = light_min) + geom_point(aes(col = status)) + facet_wrap(~ exp) + # geom_smooth() 
  geom_hline(yintercept = 0) + geom_vline(xintercept = 0) +
  scale_x_continuous(limits = c(-0.5, 4)) + scale_y_continuous(limits = c(-3.5, 2.7))
```

```{r}
test2 %>% ggplot() +
  aes(x = nutrient_max, y = light_max) + geom_point(aes(col = status)) +  facet_wrap(~ exp) + #geom_smooth() +
  geom_hline(yintercept = 0) + geom_vline(xintercept = 0) +
  scale_x_continuous(limits = c(-0.5, 4)) + scale_y_continuous(limits = c(-3.5, 2.7))
```


```{r}
cor.test(test2$light_max, test2$nutrient_max) # marginally significant
cor.test(test2$light_min, test2$nutrient_min)
cor.test(test2$light_max, test2$light_min)
cor.test(test2$nutrient_min, test2$nutrient_max)
```

```{r}
save(file = 'results.RData', list = ls())
```

