---
title: "Resource competition"
author: "ZZ"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    number_sections: true
    toc_float:
      collapsed: no
      smooth_scroll: no
---
\


**The question**:\
Here, we aim to test whether **resource availability** (light and/or nutrient) affects the competitive outcomes between **alien** and **native** plants.

**The design**:\
In Konstanz, Germany, Angelina grew 16 species (8 aliens and 8 natives), either alone, or with intraspecific competition, or with interspecific competition (32 out of the 64 alien-native pairs were included). \

In Changchun, China, Yanjie's group did a similar experiment with 8 species (4 aliens and 4 natives). They also included alien-alien and native-native pairs.\


**The analyses**:\
First, I pooled two experiments together, with the alien-alien and native-native pairs in Yanjie's experiment removed (model: m0).\
Then, I analyzed the two experiments seperately (models: m_zz & m_yj).\

**The results**:\
Most importantly, all three models showed that aliens were more competitive than natives, when both nutrient and light availabilities were high. This is indicated by the significant, or margnially significant, interactions between status (alien or native), fertilization and shading.


**To do**:\
1. The trade-off between light competition and nutrient competition.


<style type="text/css">

body{ /* Normal  */
      font-size: 16px;
      font-family: "Times New Roman"
  }
td {  /* Table  */
  font-size: 14px;
}
h1.title {
  font-size: 34px;
  color: Black;
  font-family: "Arial";
}
h1 { /* Header 1 */
  font-size: 26px;
  color: DarkBlue;
  font-family: "Arial";
}
h2 { /* Header 2 */
    font-size: 22px;
    font-family: "Arial";
    color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Arial";
  color: DarkBlue;
}

h4 { /* Header 4 */
  font-size: 16px;
  font-family: "Arial";
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 12px;
}
</style>


```{r}
knitr::opts_chunk$set(message = F,warning = F)
dir.create('fig', showWarnings = F)
dir.create('table', showWarnings = F)
```


```{r}
library(tidyverse)
library(Taxonstand)
library(nlme)
library(knitr)
library(kableExtra)
library(gridExtra)
library(effects)
load('results.RData')
source('functions.R')
```

# read data
```{r}
dat_yj_raw <- read.csv('Data_ABcomp_HFJ.csv')
dat_zz <- read.csv('Konstanz20210318.csv') %>% 
  # dplyr::select(- Box, -Box.2)
  rename(cage_rand1 = Box,
         cage_rand2 = Box.2)
colnames(dat_zz) <- colnames(dat_zz) %>% tolower()


dat_yj <- dat_yj_raw %>% 
  dplyr::select(Nr., Target_Species, Target_Status, Competition, Competitor_Species, Competitor_Status,
         N_Treatment, Shading_Treatment,
         Target_Shoot, Competitor_Shoot, Target_Root)

```


```{r}
# make pots with two data points into two rows
dat_yj_t <- dat_yj %>% # target individuals
  dplyr::select(-Competitor_Shoot) %>% 
  rename(biomass = Target_Shoot,
         target  = Target_Species,
         status  = Target_Status,
         comp        = Competitor_Species,
         status_comp = Competitor_Status,
         comp_treat  = Competition ,
         fert  = N_Treatment,
         light =  Shading_Treatment,
         id = Nr., 
         root = Target_Root)

dat_yj_c <- dat_yj %>% # competitors
  dplyr::select(-Target_Shoot) %>% 
  rename(biomass = Competitor_Shoot,
         target  = Competitor_Species,
         status  = Competitor_Status,
         comp        = Target_Species,
         status_comp = Target_Status,
         comp_treat  = Competition ,
         fert  = N_Treatment,
         light =  Shading_Treatment,
         id = Nr.,
         root = Target_Root) %>% 
  mutate(root = NA)


dat_yj <- rbind(dat_yj_t, dat_yj_c) %>% 
  filter(status != 'Without') %>% 
  mutate(status = ifelse(status == 'Self', as.character(status_comp), as.character(status))) %>% 
  #----- change names ---------#
  mutate(status      = fct_recode(status, 'alien' = 'Invasive', 'native' = 'Native'),
         status_comp = fct_recode(status_comp, 'alien' = 'Invasive', 'native' = 'Native',
                                  'intra' = 'Self', 'alone' = 'Without'),
         light      = fct_recode(light, 'high light' = 'Control', 'low light'= 'Shading'),
         fert       = tolower(fert),
         comp_treat = tolower(comp_treat),
         comp = ifelse(comp == 'No_competitor', 'Non', as.character(comp))
         ) %>% 
  #----- change reference levels -------#
  mutate(light = fct_relevel(light, 'low light'),
         fert = fct_relevel(fert, 'low')) %>% 
  droplevels()

# remove pots contained dead (NA) plants
list <- dat_yj$id[is.na(dat_yj$biomass)] %>% unique
dat_yj <- dat_yj %>% 
  filter(!id %in% list)
```


```{r}
# standardize names
sp_list <- dat_yj$target %>% unique %>% str_replace('_', ' ')



sp_list <- TPL(sp_list) %>% 
  dplyr::select(Taxon, Family) %>% 
  mutate(Taxon = str_replace(Taxon, ' ', '_')) 

dat_yj <- dat_yj %>% 
  left_join(sp_list, by = c('target' = 'Taxon')) %>% 
  rename(family_tar = Family) %>% 
  left_join(sp_list, by = c('comp' = 'Taxon')) %>% 
  rename(family_comp = Family) %>% 
  mutate(family_comp = ifelse(is.na(family_comp), 'Non', as.character(family_comp)))

rm(sp_list, list, dat_yj_c, dat_yj_t)
```


```{r}
dat_zz <- dat_zz %>% 
  mutate(status = fct_recode(status, 'alien' = 'ex', 'native' = 'nat'),
         light  = fct_recode(light, 'high light' = 'W', 'low light' = 'G'),
         fert   = fct_recode(fert , 'high' = 'H', 'low' = 'L'),
         target = str_remove(target, '\\s\\(\\w*\\)'), # remove braces and characters between them
         comp   = str_remove(comp, '\\s\\(\\w*\\)'),
         target = str_replace(target, '\\s', '_'),
         comp   = str_replace(comp, '\\s', '_')
         ) %>% 
  #----- change reference levels -------#
  mutate(light = fct_relevel(light, 'low light'),
         fert = fct_relevel(fert, 'low')) %>% 
  droplevels()
```







# analyses
## two datasets pooled
merge the two datasets (only alone, intraspecific competition, and alien-native pairs)
```{r}
dat <- dat_yj %>% 
  filter(as.character(status) != as.character(status_comp)) %>% # exclude native-native, and alien-alien combs
  dplyr::select(-status_comp) %>% 
  bind_rows(dat_zz) %>% 
  mutate(exp = str_extract(id, '^.')) %>% # distinguish the two experiments
  mutate(cage_rand1 = ifelse(is.na(cage_rand1), as.character(light), as.character(cage_rand1)),
         cage_rand2 = ifelse(is.na(cage_rand2), as.character(light), as.character(cage_rand2))
         )

```

add contrasts\
T1: alone vs competition\
T2: intra- vs inter- competition
```{r}
mat <- matrix(c(1/3, 1/3, 1/3,  
                -1,  0.5, 0.5,
                0,     1, -1), ncol=3)

mat_solve = solve(t(mat))[,2:3] %>% 
  as.data.frame() %>% 
  mutate(comp_treat = c('alone', 'inter', 'intra')) %>% 
  rename(T_alone_comp = V1,
         T_intra_inter = V2)

dat <- dat %>% 
  left_join(mat_solve)
rm(mat)

# for random slope
dat <- dat %>% 
  mutate(light_slope = paste(target, light, sep = '_'),
         fert_slope = paste(target, fert, sep = '_'))
```


```{r, eval = F}
u <- rep(1, nrow(dat))
m0 <- lme(log(biomass) ~ status*(T_alone_comp + T_intra_inter) * fert * light, # + length_in_mm
          random=list(u = pdBlocked(list(
                                        pdIdent(form = ~ exp-1),
                                        pdIdent(form = ~ family_tar-1),
                                        pdIdent(form = ~ target-1),
                                        pdIdent(form = ~ light_slope -1),
                                        pdIdent(form = ~ fert_slope -1),  # similar to random slope
                                        pdIdent(form = ~ family_comp-1),
                                        pdIdent(form = ~ comp-1)
                                        ))),
          weights = varComb(varIdent(form= ~1|target),
                            varIdent(form= ~1|comp_treat)),
          control = lmeControl(msMaxIter = 10000, msMaxEval = 10000),
          data=dat)


```


```{r}
m0 %>% Ftable
```



## seperately
### Germany (Konstanz, ZZ)
```{r}
dat_zz_only <- dat %>% 
  filter(exp == 'A')


```



```{r, eval = F}
u <- rep(1, nrow(dat_zz_only))
m_zz <- lme(log(biomass) ~ status*(T_alone_comp + T_intra_inter) * fert * light, 
            random=list(u = pdBlocked(list(
                                          pdIdent(form = ~ family_tar-1),
                                          pdIdent(form = ~ target-1),
                                          pdIdent(form = ~ family_comp-1),
                                          pdIdent(form = ~ light_slope-1),
                                          pdIdent(form = ~ fert_slope-1),
                                          pdIdent(form = ~ comp-1)
                                          ))), # adding pot id or block does not improve the model, but consume lots of computation.
            weights = varComb(varIdent(form= ~1|target),
                              varIdent(form= ~1|comp_treat)
                              ),
            control = lmeControl(msMaxIter = 100000),
            data=dat_zz_only)

```


```{r}
m_zz %>% Ftable
```



### China (Changchun, YJ)
As there is competition between aliens and between natives, another contrast - alien vs native competitor - was added. 
```{r}
mat <- matrix(c(1/4, 1/4, 1/4, 1/4,  
                -1,  1/3, 1/3, 1/3, # alone vs comp
                0,   1,   -0.5, -0.5,  # 
                0,   0,   1,  -1), # native vs alien
              ncol = 4)

mat_solve_yj = solve(t(mat))[,2:4] %>% 
  as.data.frame() %>% 
  mutate(status_comp = c('alone', 'intra', 'alien', 'native')) %>% 
  rename(T_alone_comp   = V1,
         T_intra_inter  = V2,
         T_alien_native = V3
         )

dat_yj_only <- dat_yj %>% 
  left_join(mat_solve_yj) %>% 
  mutate(light_slope = paste(target, light, sep = '_'),
         fert_slope = paste(target, fert, sep = '_'))
rm(mat)
```


```{r, eval = F}
u <- rep(1, nrow(dat_yj_only))
m_yj <- lme(log(biomass) ~ status*(T_alone_comp + T_intra_inter + T_alien_native) * fert * light, # + length_in_mm
            random=list(u = pdBlocked(list(
                                          pdIdent(form = ~ family_tar-1),
                                          pdIdent(form = ~ target-1),
                                          pdIdent(form = ~ family_comp-1),
                                          pdIdent(form = ~ light_slope-1),
                                          pdIdent(form = ~ fert_slope-1),
                                          pdIdent(form = ~ comp-1)
                                          ))),
          weights = varComb(varIdent(form= ~1|target),
                            varIdent(form= ~1|status_comp) # removing this is ok
                            ),
          control = lmeControl(msMaxIter = 10000),
          data=dat_yj_only)

```



```{r}
m_yj %>% Ftable
```


Export all tables in one file

```{r}
tables <- lapply(list(m_yj, m_zz, m0), Ftable_ext)
table_in_one <- tables[[1]] %>% # China
  left_join(tables[[2]], by = 'items') %>% # Germany 
  full_join(tables[[3]], by = 'items') %>%  # both
  replace_na(list(`.y` = ' ',
                  ` ` = ' '
                  )) %>% 
  mutate_if(is.numeric, round,3)
  

# table_in_one %>% kable() %>% 
#   kable_styling()
```


```{r}
write.csv(table_in_one, 'table/table_in_one.csv', fileEncoding = "UTF-8")
```




# Figures for analysis I
All figures are ploted with modelled values and back-transformed to raw scale.\


```{r}
theme <-theme(panel.background = element_blank(),
              panel.border     = element_rect(fill = NA), 
              strip.background = element_blank(),
              strip.text.x = element_text(size = 10),
              axis.text.x = element_text(size = 8, angle = 0, hjust = 0.5, colour = 'black'),
              axis.text.y  = element_text(size = 8, colour = 'black'),
              axis.title = element_text(size = 10),# size of x  y lab
              #legend.position   = c(0.9, 0.85),
              legend.title      = element_text(size = 8),
              legend.text       = element_text(size = 6),
              legend.key.height = unit(0.8,'line'),
              legend.background = element_blank(),
              legend.key        = element_rect(colour = NA, fill = NA),
              plot.tag    = element_text(size = 9.5, face  = 'bold'),
              plot.title  = element_text(hjust = 0.5, size = 10),
              plot.margin = margin(t = 0.5, r = 0.1, b = 0, l = 0.1, unit = "cm"))
```



modelled values\

```{r, eval = F}
u <- rep(1, nrow(dat)) # for pooled data
m_resource  <- update(m0, .~ status * fert * light) # fig. 1
m_comp <- update(m0, ~ light * fert* comp_treat) # fig. 2
```


percentage change
```{r}
# # fert
# test  <- update(m0, .~ fert) # fig. 1
# test <- effect('fert', test) %>% as.data.frame() %>% dplyr::select(fit) %>% exp
# print(paste('fert = ',test[2,1]/test[1,1] - 1))
# 
# # light
# test  <- update(m0, .~ light) # fig. 1
# test <- effect('light', test) %>% as.data.frame() %>% dplyr::select(fit) %>% exp
# print(paste('light = ',test[2,1]/test[1,1] - 1))
# 
# 
# # fert:light
# test  <- update(m0, .~ fert*light) # fig. 1
# test <- effect('fert:light', test) %>% as.data.frame() %>% dplyr::select(fit) %>% exp
# print(paste('fert:light = ',test[4,1]/test[1,1] - 1))
# 
# # competition
# test  <- update(m0, .~ T_alone_comp + T_intra_inter) # fig. 1
# effect('T_alone_comp', test, xlevels = list(T_alone_comp = c(-0.66666667, 0.3333333))) %>% as.data.frame() %>% dplyr::select(fit) %>% exp
# print(paste('comp = ',test[2,1]/test[1,1] - 1))
# 
# #intra_inter
# test  <- update(m0, .~ comp_treat) # fig. 1
# test <- effect('comp_treat', test) %>% as.data.frame() %>% dplyr::select(fit) %>% exp
# test[2,1]/test[3,1]-1
# 
# # main inter
# test <- effect('status:fert:light', m_resource) %>% as.data.frame() %>% 
#   mutate(fit = exp(fit))
# 
# test[7,4]/test[8,4]-1 # both high
# test[5,4]/test[6,4]-1 # high light
# test[3,4]/test[4,4]-1 # high nutrient
# test[1,4]/test[2,4]-1 # both low

```



```{r, eval = F}
u <- rep(1, nrow(dat_yj_only)) # China
m_resource_yj <- update(m_yj, .~ status * fert * light)
m_comp_resource_yj <- update(m_yj, .~ status * comp_treat * light)
m_status <- update(m_yj, .~ status * status_comp)
```

```{r, eval = F}
u <- rep(1, nrow(dat_zz_only))
m_resource_zz <- update(m_zz, .~ status * fert * light)
m_comp_resource_zz <- update(m_zz, .~ status * comp_treat * light)
```


## origin:fert:light

```{r}
#function
plot_resource <- function(model = m_resource, ymax = 17, title = NA){
  t <- effect('status:fert:light', model) %>% as.data.frame() %>% 
    mutate(xaxis = c(1, 2, 4, 5, 7, 8, 10, 11))
  plot <- t %>% 
    mutate(origin = status) %>% # rename
    ggplot() +
    geom_point(aes(x = xaxis, y = fit %>% exp, col = origin),size = 2.5) +
    geom_errorbar(aes(ymin = lower %>% exp, ymax = upper %>% exp, x = xaxis, col = origin), 
                  width = 0,  size = 0.6,
                  inherit.aes = FALSE) +
    scale_x_continuous(limits = c(0.5, 11.5), breaks = c(1.5, 4.5, 7.5, 10.5), 
                       labels = c("Low resources", "+ nutrient", '+ light', '+ nutrient & light')) + 
    scale_y_continuous(limits = c(0, ymax)) + 
    labs(x = '', y = 'Aboveground biomass [g]\n', col = 'Origin') + theme
  if (!is.na(title)) {
    plot <- plot + labs(title = title)
  }
  return(plot)
}
```

### joint (Fig. 1)

```{r}
p1 <- plot_resource()
```


```{r}
ggsave('./fig/pooled.pdf', p1, height = 7/2.54, width = 11/2.54, useDingbats=FALSE)
```


### separate (Fig. s1)
```{r}

p_s1a <- plot_resource(m_resource_yj, ymax = 50, title = 'China\n' )   + labs(tag = 'a') + 
  theme(axis.text.x = element_text(size = 7),
        axis.text.y  = element_text(size = 7))
p_s1b <- plot_resource(m_resource_zz, ymax = 4,  title = 'Germany\n' ) + labs(tag = 'b') + 
  theme(axis.text.x = element_text(size = 7),
        axis.text.y  = element_text(size = 7))

ggsave('./fig/s1.pdf', grid.arrange(p_s1a, p_s1b, ncol =2, nrow = 1), 
       height = 8.5/2.54, width = 20/2.54, useDingbats=FALSE)

```



## fert:light:competition (Fig. 2)
```{r}
t_comp <- effect('light * fert* comp_treat', m_comp) %>% as.data.frame() %>% 
  arrange(desc(light), desc(fert)) %>% 
  mutate(xaxis = c(1:3, 5:7, 9:11, 13:15))
```

```{r}
col_fig <- c('#F39C12', '#115888', '#85C1E9')

p2 <- t_comp %>% 
  ggplot() +
  geom_point(aes(x = xaxis, y = fit %>% exp, col = comp_treat),size = 2.5) +
  geom_errorbar(aes(ymin = lower %>% exp, ymax = upper %>% exp, x = xaxis, col = comp_treat), 
                width = 0,  size = 0.6,
                inherit.aes = FALSE) +
  scale_x_continuous(limits = c(0.5, 15.5), breaks = c(2, 6, 10, 14), 
                     labels = c("Low resources", "+ nutrient", '+ light', '+ nutrient & light')) + 
  scale_y_continuous(limits = c(0, 17)) + 
  scale_color_manual(values = col_fig) + 
  labs(x = '', y = 'Aboveground biomass [g]\n', col = 'Competition') + theme

```




```{r}
ggsave('./fig/comp.pdf', p2, height = 7/2.54, width = 11/2.54, useDingbats=FALSE)
```



## origin:origin_comp (Fig. s2)
```{r}
t_status <- effect('status:status_comp', m_status) %>% as.data.frame() %>% 
  arrange(status)
t_status <- t_status[c(1,4, 5, 8),] %>% 
  mutate(xaxis = c(1:2, 4:5))

```

```{r}
p_comp <- t_status %>% 
  ggplot() +
  geom_point(aes(x = xaxis, y = fit %>% exp, col = status), 
             size = 2.5) +
  geom_errorbar(aes(ymin = lower %>% exp, ymax = upper %>% exp, x = xaxis, col = status), 
                width = 0,  inherit.aes = FALSE,
                size = 0.6) +
  scale_x_continuous(limits = c(0.5, 5.5), breaks = c(1, 2, 4, 5), 
                     labels = c('Alien', 'Native', 'Alien', 'Native')) + 
  scale_y_continuous(limits = c(0, 20)) + 
  labs(x = '\nOrigin of the competitor', y = 'Aboveground biomass [g]\n', 
       col = 'Origin of the target', title = 'China')+ theme
p_comp
```


```{r}
ggsave('./fig/origin.pdf', p_comp, height = 7/2.54, width = 11/2.54, useDingbats=FALSE)
```

## origin:competition:light (Fig. S3)


```{r}
t_light_comp_zz <- effect('status:comp_treat:light', m_comp_resource_zz) %>% as.data.frame() %>% 
  filter(comp_treat != 'alone') %>% 
  arrange(comp_treat) %>% 
  mutate(xaxis = c(1, 2, 4, 5, 1, 2, 4, 5))




p_light_comp_zz <- t_light_comp_zz %>% 
  mutate(comp_treat = ifelse(comp_treat == 'intra', 'Intraspecific competition', 'Interspecific competition')) %>% 
  ggplot() +
  geom_point(aes(x = xaxis, y = fit %>% exp, col = status), size = 2.5) +
  geom_errorbar(aes(ymin = lower %>% exp, ymax = upper %>% exp, x = xaxis, col = status), 
                width = 0,  size = 0.6,
                inherit.aes = FALSE) +
  scale_x_continuous(limits = c(0.5, 5.5), breaks = c(1.5, 4.5), 
                     labels = c('Low light-intensity', 'High light-intensity')) + 
  scale_y_continuous(limits = c(0, 4)) + facet_wrap( ~ comp_treat) + 
  # scale_color_manual(values = col_fig) + 
  labs(x = '', y = 'Aboveground biomass [g]\n', col = 'Origin', title = 'Germany') + theme +
  theme(panel.spacing = unit(2, "lines"),
        axis.text.x = element_text(size = 7),
        axis.text.y  = element_text(size = 7))

ggsave('./fig/origin_light_comp.pdf', p_light_comp_zz, height = 7/2.54, width = 14/2.54, useDingbats=FALSE)
```



```{r}
# save(file = 'results.RData', list = ls())
```

