---
title: "resource_comp"
author: "ZZ"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    number_sections: true
    toc_float:
      collapsed: no
      smooth_scroll: no
---




**The question**:
Here, we aim to test whether **resource availability** (light and/or nutrient) affects the competitive outcomes between **alien** and **native** plants.

**The design**:\
Angelina grew 16 species (8 aliens and 8 natives), either alone, with intraspecific competition, or with interspecific competition (only alien-native pairs, and included 32 pairs). The experiment was conducted in Konstanz, Germany\

Yanjie's group did a similar experiment with 8 species (4 aliens and 4 natives), in Changchun, China. But they also included alien-alien and native-native pairs.\


**The analyses**:\
First, I pooled two experiments together, removing the alien-alien and native-native pairs in Yanjie's experiment (model: m0).\
Then, I analyzed the two experiments seperately (models: m_zz & m_yj).\

**The results**:\
Most importantly, all three models showed that aliens were more competitive than natives, when both nutrient and light availabilities are high. This is indicated by the significant, or margnially significant, interactions between status (alien or native), fertilization and shading.



<style type="text/css">

body{ /* Normal  */
      font-size: 16px;
      font-family: "Times New Roman"
  }
td {  /* Table  */
  font-size: 14px;
}
h1.title {
  font-size: 34px;
  color: Black;
  font-family: "Arial";
}
h1 { /* Header 1 */
  font-size: 26px;
  color: DarkBlue;
  font-family: "Arial";
}
h2 { /* Header 2 */
    font-size: 22px;
    font-family: "Arial";
    color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Arial";
  color: DarkBlue;
}

h4 { /* Header 4 */
  font-size: 16px;
  font-family: "Arial";
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 12px;
}
</style>


```{r}
knitr::opts_chunk$set(message = F,warning = F)
```


```{r}
library(tidyverse)
library(Taxonstand)
library(nlme)
library(knitr)
library(kableExtra)
load('results.RData')
source('functions.R')
```

# read data
```{r}
dat_yj_raw <- read.csv('Data_ABcomp_HFJ.csv')
dat_zz <- read.csv('Konstanz.csv')
colnames(dat_zz) <- colnames(dat_zz) %>% tolower()

dat_yj <- dat_yj_raw %>% 
  dplyr::select(Nr., Target_Species, Target_Status, Competition, Competitor_Species, Competitor_Status,
         N_Treatment, Shading_Treatment,
         Target_Shoot, Competitor_Shoot)

```


```{r}
# make pots with two data points into two rows
dat_yj_t <- dat_yj %>% # target individuals
  dplyr::select(-Competitor_Shoot) %>% 
  rename(biomass = Target_Shoot,
         target  = Target_Species,
         status  = Target_Status,
         comp        = Competitor_Species,
         status_comp = Competitor_Status,
         comp_treat  = Competition ,
         fert  = N_Treatment,
         light =  Shading_Treatment,
         id = Nr.)

dat_yj_c <- dat_yj %>% # competitors
  dplyr::select(-Target_Shoot) %>% 
  rename(biomass = Competitor_Shoot,
         target  = Competitor_Species,
         status  = Competitor_Status,
         comp        = Target_Species,
         status_comp = Target_Status,
         comp_treat  = Competition ,
         fert  = N_Treatment,
         light =  Shading_Treatment,
         id = Nr.)
dat_yj <- rbind(dat_yj_t, dat_yj_c) %>% 
  filter(status != 'Without') %>% 
  mutate(status = ifelse(status == 'Self', as.character(status_comp), as.character(status))#,
         #status_comp = ifelse(status_comp == 'Self', as.character(status), as.character(status_comp)),
         ) %>% 
  #----- change names ---------#
  mutate(status      = fct_recode(status, 'alien' = 'Invasive', 'native' = 'Native'),
         status_comp = fct_recode(status_comp, 'alien' = 'Invasive', 'native' = 'Native',
                                  'intra' = 'Self', 'alone' = 'Without'),
         light      = fct_recode(light, 'Ambient' = 'Control'),
         comp_treat = tolower(comp_treat),
         comp = ifelse(comp == 'No_competitor', 'Non', as.character(comp))
         ) %>% 
  #----- change reference levels -------#
  mutate(light = fct_relevel(light, 'Shading'),
         fert = fct_relevel(fert, 'Low')) %>% 
  droplevels()

# remove pots contained dead (NA) plants
list <- dat_yj$id[is.na(dat_yj$biomass)] %>% unique
dat_yj <- dat_yj %>% 
  filter(!id %in% list)

sp_list <-dat_yj$target %>% unique %>% str_replace('_', ' ')
sp_list <- TPL(sp_list) %>% 
  dplyr::select(Taxon, Family) %>% 
  mutate(Taxon = str_replace(Taxon, ' ', '_')) 
dat_yj <- dat_yj %>% 
  left_join(sp_list, by = c('target' = 'Taxon')) %>% 
  rename(family_tar = Family) %>% 
  left_join(sp_list, by = c('comp' = 'Taxon')) %>% 
  rename(family_comp = Family) %>% 
  mutate(family_comp = ifelse(is.na(family_comp), 'Non', as.character(family_comp)))
rm(sp_list, list, dat_yj_c, dat_yj_t)
```


```{r}
dat_zz <- dat_zz %>% 
  mutate(status = fct_recode(status, 'alien' = 'ex', 'native' = 'nat'),
         light  = fct_recode(light, 'Ambient' = 'W', 'Shading' = 'G'),
         fert   = fct_recode(fert , 'High' = 'H', 'Low' = 'L'),
         target = str_remove(target, '\\s\\(\\w*\\)'), # remove braces and characters between them
         comp   = str_remove(comp, '\\s\\(\\w*\\)'),
         target = str_replace(target, '\\s', '_'),
         comp   = str_replace(comp, '\\s', '_')
         ) %>% 
  #----- change reference levels -------#
  mutate(light = fct_relevel(light, 'Shading'),
         fert = fct_relevel(fert, 'Low')) %>% 
  droplevels()
```







# analyses

## only native-alien, two dataset pooled
merge two datasets
```{r}
dat <- dat_yj %>% 
  filter(as.character(status) != as.character(status_comp)) %>% # exclude native-native, and alien-alien combs
  dplyr::select(-status_comp) %>% 
  bind_rows(dat_zz) %>% 
  mutate(exp = str_extract(id, '^.')) # distinguish the two experiments
```

add contrast\
T1: alone vs competition\
T2: intra- vs inter- competition
```{r}
mat <- matrix(c(1/3, 1/3, 1/3,  
                -1,  0.5, 0.5,
                0,     1, -1), ncol=3)

mat_solve = solve(t(mat))[,2:3] %>% 
  as.data.frame() %>% 
  mutate(comp_treat = c('alone', 'inter', 'intra')) %>% 
  rename(T_alone_comp = V1,
         T_intra_inter = V2)
dat <- dat %>% 
  left_join(mat_solve)
rm(mat)
```


```{r, eval = F}
u <- rep(1, nrow(dat))
dat <- dat %>% 
  mutate(light_slope = paste(target, light, sep = '_'),
         fert_slope = paste(target, fert, sep = '_'))
m0 <- lme(log(biomass) ~ status*(T_alone_comp + T_intra_inter) * fert * light, # + length_in_mm
          random=list(u = pdBlocked(list(
                                        pdIdent(form = ~ exp-1),
                                        pdIdent(form = ~ family_tar-1),
                                        pdIdent(form = ~ target-1),
                                        pdIdent(form = ~ light_slope -1),
                                        pdIdent(form = ~ fert_slope -1),  # similar to random slope
                                        pdIdent(form = ~ family_comp-1),
                                        pdIdent(form = ~ comp-1)
                                        ))),
          weights = varComb(varIdent(form= ~1|target),
                            varIdent(form= ~1|comp_treat)),
          control = lmeControl(msMaxIter = 10000, msMaxEval = 10000),
          data=dat)
```


```{r}
m0 %>% Ftable
```



## seperately
### Germany (Konstanz, ZZ)
```{r}
dat_zz_only <- dat %>% 
  filter(exp == 'A')
```



```{r, eval = F}
u <- rep(1, nrow(dat_zz_only))
m_zz <- lme(log(biomass) ~ status*(T_alone_comp + T_intra_inter) * fert * light, 
            random=list(u = pdBlocked(list(
                                          pdIdent(form = ~ family_tar-1),
                                          pdIdent(form = ~ target-1),
                                          pdIdent(form = ~ family_comp-1),
                                          pdIdent(form = ~ light_slope-1),
                                          pdIdent(form = ~ fert_slope-1),
                                          pdIdent(form = ~ comp-1)
                                          ))), # adding pot id does not improve the model, but consume lots of computation.
            weights = varComb(varIdent(form= ~1|target),
                              varIdent(form= ~1|comp_treat)
                              ),
            control = lmeControl(msMaxIter = 100000),
            data=dat_zz_only)

test2 <- update(m_zz, .~ status:fert:light - 1)
```


```{r}
m_zz %>% Ftable
```



### China (Changchun, YJ)
As there is competition between aliens and between natives, another contrast - alien vs native competitor - was added. 
```{r}
mat <- matrix(c(1/4, 1/4, 1/4, 1/4,  
                -1,  1/3, 1/3, 1/3, # alone vs comp
                0,   1,   -0.5, -0.5,  # 
                0,   0,   1,  -1), # native vs alien
              ncol = 4)

mat_solve_yj = solve(t(mat))[,2:4] %>% 
  as.data.frame() %>% 
  mutate(status_comp = c('alone', 'intra', 'alien', 'native')) %>% 
  rename(T_alone_comp   = V1,
         T_intra_inter  = V2,
         T_alien_native = V3
         )
dat_yj_only <- dat_yj %>% 
  left_join(mat_solve_yj) %>% 
  mutate(light_slope = paste(target, light, sep = '_'),
         fert_slope = paste(target, fert, sep = '_'))
rm(mat)
```


```{r, eval = F}
u <- rep(1, nrow(dat_yj_only))
m_yj <- lme(log(biomass) ~ status*(T_alone_comp + T_intra_inter + T_alien_native) * fert * light, # + length_in_mm
            random=list(u = pdBlocked(list(
                                          pdIdent(form = ~ family_tar-1),
                                          pdIdent(form = ~ target-1),
                                          pdIdent(form = ~ family_comp-1),
                                          pdIdent(form = ~ light_slope-1),
                                          pdIdent(form = ~ fert_slope-1),
                                          pdIdent(form = ~ comp-1)
                                          ))),
          weights = varComb(varIdent(form= ~1|target),
                            varIdent(form= ~1|status_comp) # removing this is ok
                            ),
          control = lmeControl(msMaxIter = 10000),
          data=dat_yj_only)

```



```{r}
m_yj %>% Ftable
```


# Figures
All figures are ploted with raw data.\

## status:fertilization:shading
Aliens are more competitive than native when both nutrient and light levels are high.\


In Yanjie's experiment, the boxplot based on raw data shows that aliens were less competitive than natives, when nutrient level is low and light level is high. This is not consistent with the other two figures. But if we use modelled values, the pattern will be consistent with the other two figures.\

So, I would prefer to plot with modelled values.
```{r, fig.width = 5, fig.height = 3.5}
theme <-theme(panel.background = element_blank(),
              panel.border     = element_rect(fill = NA), 
              strip.background = element_blank(),
              strip.text.x = element_text(size = 10),
              axis.text.x = element_text(size = 8, angle = 0, hjust = 0.5, colour = 'black'),
              axis.text.y  = element_text(size = 8, colour = 'black'),
              axis.title = element_text(size = 10),# size of y lab
              #legend.position   = c(0.9, 0.85),
              legend.title      = element_text(size = 10),
              legend.text       = element_text(size = 8),
              legend.key.height = unit(0.7,'line'),
              legend.background = element_blank(),
              legend.key        = element_rect(colour = NA, fill = NA),
              plot.tag    = element_text(size = 8, face  = 'bold'),
              plot.title  = element_text(hjust = 0.5, size = 12),
              plot.margin = margin(t = 0.5, r = 0.1, b = 0, l = 0.1, unit = "cm"))

dat %>% 
  ggplot(aes(x = fert, y = biomass, fill = status)) + geom_boxplot() +
  facet_wrap( ~ light) + labs(title = 'Pooled', y = 'Aboveground biomass\n', x = '\nNutrient level') + theme

dat_zz_only %>% 
  ggplot(aes(x = fert, y = biomass, fill = status)) + geom_boxplot() +
  facet_wrap( ~ light) + labs(title = 'Pooled', y = 'Aboveground biomass\n', x = '\nNutrient level') + theme


dat_yj_only %>% 
  ggplot(aes(x = fert, y = biomass, fill = status)) + geom_boxplot() +
  facet_wrap( ~ light) + labs(title = 'Pooled', y = 'Aboveground biomass\n', x = '\nNutrient level') + theme
```




## T_alone_comp:light
```{r, fig.width = 5, fig.height = 3.5}
dat %>% 
  ggplot(aes(x = comp_treat, y = biomass)) + geom_boxplot() +
  facet_wrap( ~ light) + labs(title = 'Pooled', y = 'Aboveground biomass\n', x = '\nCompetition') + theme

```

## T_alone_comp:fert
```{r, fig.width = 5, fig.height = 3.5}
dat %>% 
  ggplot(aes(x = comp_treat, y = biomass)) + geom_boxplot() +
  facet_wrap( ~ fert) + labs(title = 'Pooled', y = 'Aboveground biomass\n', x = '\nCompetition') + theme
```

## fert:light
```{r, fig.width = 5, fig.height = 3.5}
dat %>% 
  ggplot(aes(x = fert, y = biomass)) + geom_boxplot() +
  facet_wrap( ~ light) + labs(title = 'Pooled', y = 'Aboveground biomass\n', x = '\nNutrient level') + theme
```


## status:T_alien_native
only significant in Yanjie's experiment. Aliens were less limited by natives than by aliens, but this does not hold for natives.
```{r, fig.width = 5, fig.height = 3.5}
dat_yj_only %>% 
  filter(status_comp %in% c('alien', 'native')) %>% 
  ggplot(aes(x = status_comp, y = biomass, fill = status)) + geom_boxplot() + theme
```


export tables
```{r}
dir.create('table', showWarnings = F)
m0 %>% Ftable %>% save_kable('table/pooled.png')
m_zz %>% Ftable %>% save_kable('table/zz_single.png')
m_yj %>% Ftable %>% save_kable('table/yj_single.png')

```

```{r}
save(file = 'results.RData', list = ls())
```

