---
title: "Resource competition"
author: "ZZ"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    number_sections: true
    toc_float:
      collapsed: no
      smooth_scroll: no
---



**The question**:\
Here, we aim to test whether **resource availability** (light and/or nutrient) affects the competitive outcomes between **alien** and **native** plants.

**The design**:\
In Konstanz, Germany, Angelina grew 16 species (8 aliens and 8 natives), either alone, or with intraspecific competition, or with interspecific competition (32 out of the 64 alien-native pairs were included). \

In Changchun, China, Huifei (Yanjie's group) grew 8 species (4 aliens and 4 natives) in a similar design with additional pairs of alien-alien and native-native.\


**The analyses**:\
First, I pooled two experiments together, with the alien-alien and native-native pairs in China removed (model: m0).\
Then, I analyzed the two experiments seperately (models: m_zz & m_yj).\

**The results**:\
Most importantly, all three models showed that aliens were more competitive than natives with co-addition of nutrients and light. This is indicated by the significant, or margnially significant, interactions between status (alien or native), fertilization and shading.



<style type='text/css'>

body{ /* Normal  */
      font-size: 16px;
      font-family: 'Times New Roman'
  }
td {  /* Table  */
  font-size: 14px;
}
h1.title {
  font-size: 34px;
  color: Black;
  font-family: 'Arial';
}
h1 { /* Header 1 */
  font-size: 26px;
  color: DarkBlue;
  font-family: 'Arial';
}
h2 { /* Header 2 */
    font-size: 22px;
    font-family: 'Arial';
    color: DarkBlue;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: 'Arial';
  color: DarkBlue;
}

h4 { /* Header 4 */
  font-size: 16px;
  font-family: 'Arial';
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 12px;
}
</style>


```{r}
knitr::opts_chunk$set(message = F,warning = F)
dir.create('fig', showWarnings = F)
dir.create('table', showWarnings = F)
```


I have saved model results in 'results20221004.RData', so one need not to re-run them.
```{r}
library(tidyverse)
library(Taxonstand)
library(nlme)
library(car)
library(knitr)
library(kableExtra)
library(gridExtra)
library(effects)
load('results20221004.RData')
source('functions.R')
```

# read data
_yj stands for Yanjie, _zz stands for Zhijie\

**variables**:\
*id*: id of pots\
*target*: species name of target species\
*comp*:  species name of competitor species\
*status*: origin of target species\
*status_comp*: origin of competitor species\
*fert*: nutrient treatment\
*light*: light treatment\
*family_tar*: family of target species\
*family_comp*: family of competitor species\
*ini_height*: initial height of target species

```{r}
names(dat_yj)
dat_yj <- read.csv('d_China.csv')
dat_zz <- read.csv('d_Germany.csv')
```


# analyses
## two datasets pooled
merge the two datasets (only alone, intraspecific competition, and alien-native pairs)
```{r, eval = F}
dat <- dat_yj %>% 
  filter(as.character(status) != as.character(status_comp)) %>% # exclude native-native, and alien-alien combs
  dplyr::select(-status_comp) %>% 
  bind_rows(dat_zz) %>% 
  mutate(exp = str_extract(id, '^.')) %>% # distinguish the two experiments
  mutate(cage_rand1 = ifelse(is.na(cage_rand1), as.character(light), as.character(cage_rand1)),
         cage_rand2 = ifelse(is.na(cage_rand2), as.character(light), as.character(cage_rand2))
         )

```

add contrasts\
T1: alone vs competition\
T2: intra- vs inter- competition
```{r, eval = F}
mat <- matrix(c(1/3, 1/3, 1/3,  
                -1,  0.5, 0.5,
                0,     1, -1), ncol=3)

mat_solve = solve(t(mat))[,2:3] %>% 
  as.data.frame() %>% 
  mutate(comp_treat = c('alone', 'inter', 'intra')) %>% 
  rename(T_alone_comp = V1,
         T_intra_inter = V2)

dat <- dat %>% 
  left_join(mat_solve)
rm(mat)

# for random slope
dat <- dat %>% 
  mutate(light_slope = paste(target, light, sep = '_'),
         fert_slope = paste(target, fert, sep = '_'))
```




```{r, eval = F}
u <- rep(1, nrow(dat))
m0 <- lme(log(biomass) ~ status*(T_alone_comp + T_intra_inter) * fert * light, 
          random=list(u = pdBlocked(list(
                                        pdIdent(form = ~ exp-1),
                                        pdIdent(form = ~ family_tar-1),
                                        pdIdent(form = ~ target-1),
                                        pdIdent(form = ~ light_slope -1), # similar to random slope
                                        pdIdent(form = ~ fert_slope -1),  # similar to random slope
                                        pdIdent(form = ~ family_comp-1),
                                        pdIdent(form = ~ comp-1),
                                        pdIdent(form = ~ cage_rand1-1),
                                        pdIdent(form = ~ cage_rand2-1),
                                        pdIdent(form = ~ id-1)
                                        ))),
          weights = varComb(varIdent(form= ~1|target),
                            varIdent(form= ~1|comp_treat)),
          control = lmeControl(msMaxIter = 10000, msMaxEval = 10000),
          data=dat)


```



```{r}
m0 %>% Ftable
```



## seperately
### Germany (Konstanz, ZZ)
```{r}
dat_zz_only <- dat %>% 
  filter(exp == 'A')
```



```{r, eval = F}
u <- rep(1, nrow(dat_zz_only))
m_zz <- lme(log(biomass) ~ status*(T_alone_comp + T_intra_inter) * fert * light, 
            random=list(u = pdBlocked(list(
                                          pdIdent(form = ~ family_tar-1),
                                          pdIdent(form = ~ target-1),
                                          pdIdent(form = ~ family_comp-1),
                                          pdIdent(form = ~ light_slope-1),
                                          pdIdent(form = ~ fert_slope-1),
                                          pdIdent(form = ~ comp-1),
                                          pdIdent(form = ~ cage_rand1-1),
                                          pdIdent(form = ~ cage_rand2-1),
                                          pdIdent(form = ~ id-1)
                                          ))), # adding pot id or block does not improve the model, but consume lots of computation.
            weights = varComb(varIdent(form= ~1|target),
                              varIdent(form= ~1|comp_treat)
                              ),
            control = lmeControl(msMaxIter = 100000),
            data=dat_zz_only)
```


```{r}
m_zz %>% Ftable
```



### China (Changchun, YJ)
As there is competition between aliens and between natives, another contrast - alien vs native competitor - was added. 
```{r}
mat <- matrix(c(1/4, 1/4, 1/4, 1/4,  
                -1,  1/3, 1/3, 1/3, # alone vs comp
                0,   1,   -0.5, -0.5,  # 
                0,   0,   1,  -1), # native vs alien
              ncol = 4)

mat_solve_yj = solve(t(mat))[,2:4] %>% 
  as.data.frame() %>% 
  mutate(status_comp = c('alone', 'intra', 'alien', 'native')) %>% 
  rename(T_alone_comp   = V1,
         T_intra_inter  = V2,
         T_alien_native = V3
         )

dat_yj_only <- dat_yj %>% 
  left_join(mat_solve_yj) %>% 
  mutate(light_slope = paste(target, light, sep = '_'),
         fert_slope = paste(target, fert, sep = '_'))
rm(mat)
```


```{r, eval = F}
u <- rep(1, nrow(dat_yj_only))
m_yj <- lme(log(biomass) ~ status*(T_alone_comp + T_intra_inter + T_alien_native) * fert * light, # + length_in_mm
            random=list(u = pdBlocked(list(
                                          pdIdent(form = ~ family_tar-1),
                                          pdIdent(form = ~ target-1),
                                          pdIdent(form = ~ family_comp-1),
                                          pdIdent(form = ~ light_slope-1),
                                          pdIdent(form = ~ fert_slope-1),
                                          pdIdent(form = ~ comp-1),
                                          pdIdent(form = ~ id-1)
                                          ))),
          weights = varComb(varIdent(form= ~1|target),
                            varIdent(form= ~1|status_comp) # removing this is ok
                            ),
          control = lmeControl(msMaxIter = 10000),
          data=dat_yj_only)

```

add initial height as the covariate
```{r, eval = F}
m_yj_co <- update(m_yj, .~ log(ini_height) + status*(T_alone_comp + T_intra_inter + T_alien_native) * fert * light)
```

without covariate (initial height)
```{r}
m_yj %>% Ftable
```

with covariate
```{r}
m_yj_co %>% Ftable
```


Export all tables in one file




```{r}
tables <- lapply(list(m_yj, m_zz, m0), Ftable_ext)
table_in_one <- tables[[1]] %>% # China
  full_join(tables[[2]], by = 'items') %>% # Germany 
  full_join(tables[[3]], by = 'items') %>%  # both
  replace_na(list(`.y` = ' ',
                  ` ` = ' '
                  )) %>% 
  mutate_if(is.numeric, round,3)
```

```{r}
write.csv(Ftable_ext(m_yj_co) %>% mutate_if(is.numeric, round,3), 'table/table_co.csv', fileEncoding = 'UTF-8')
```

```{r}
# table_in_one %>% kable() %>% 
#   kable_styling()
```


```{r}
write.csv(table_in_one, 'table/table_in_one20221004.csv', fileEncoding = 'UTF-8')
```



# Figures 
All figures are plotted with modelled values and back-transformed to raw scale.\

```{r}
theme <-  theme(panel.background = element_blank(),
                panel.border     = element_rect(fill = NA), 
                strip.background = element_blank(),
                strip.text.x = element_text(size = 8),
                axis.text.x = element_text(size = 7, angle = 0, hjust = 0.5, colour = 'black'),
                axis.text.y  = element_text(size = 7, colour = 'black'),
                axis.title = element_text(size = 8),# size of x  y lab
                #legend.position   = c(0.9, 0.85),
                legend.title      = element_text(size = 8),
                legend.text       = element_text(size = 7),
                legend.key.height = unit(0.8,'line'),
                legend.background = element_blank(),
                legend.key        = element_rect(colour = NA, fill = NA),
                plot.tag    = element_text(size = 8, face  = 'bold'),
                plot.title  = element_text(hjust = 0.5, size = 10),
                plot.margin = margin(t = 0.5, r = 0.1, b = 0, l = 0.1, unit = 'cm'))
```



modelled values\

```{r, eval = F}
u <- rep(1, nrow(dat)) # for pooled data
m_resource  <- update(m0, .~ status * fert * light) # fig. 1
m_comp <- update(m0, ~ light * fert* comp_treat) # fig. 2
```


percentage change

```{r, eval = F}
# random effect
random=list(u = pdBlocked(list(
                              pdIdent(form = ~ exp-1),
                              pdIdent(form = ~ family_tar-1),
                              pdIdent(form = ~ target-1),
                              pdIdent(form = ~ light_slope -1),
                              pdIdent(form = ~ fert_slope -1),  # similar to random slope
                              pdIdent(form = ~ family_comp-1),
                              pdIdent(form = ~ comp-1)
                              )))

# fert
test  <- update(m0, .~ fert, random = random) # fig. 1
test <- effect('fert', test) %>% as.data.frame() %>% dplyr::select(fit) %>% exp
print(paste('fert = ',test[2,1]/test[1,1] - 1))

# light
test  <- update(m0, .~ light, random = random) # fig. 1
test <- effect('light', test) %>% as.data.frame() %>% dplyr::select(fit) %>% exp
print(paste('light = ',test[2,1]/test[1,1] - 1))


# fert:light
test  <- update(m0, .~ fert*light, random = random) # fig. 1
test <- effect('fert:light', test) %>% as.data.frame() %>% dplyr::select(fit) %>% exp
print(paste('fert:light = ',test[4,1]/test[1,1] - 1))

# competition
test  <- update(m0, .~ T_alone_comp + T_intra_inter, random = random) # fig. 1
effect('T_alone_comp', test, xlevels = list(T_alone_comp = c(-0.66666667, 0.3333333))) %>% as.data.frame() %>% dplyr::select(fit) %>% exp
print(paste('comp = ',test[2,1]/test[1,1] - 1))

#intra_inter
test  <- update(m0, .~ comp_treat, random = random) # fig. 1
test <- effect('comp_treat', test) %>% as.data.frame() %>% dplyr::select(fit) %>% exp
test[2,1]/test[3,1]-1

# main inter
test <- effect('status:fert:light', m_resource) %>% as.data.frame() %>%
  mutate(fit = exp(fit))

test[7,4]/test[8,4]-1 # both high
test[5,4]/test[6,4]-1 # high light
test[3,4]/test[4,4]-1 # high nutrient
test[1,4]/test[2,4]-1 # both low

```



```{r, eval = F}
u <- rep(1, nrow(dat_yj_only)) # China
m_resource_yj <- update(m_yj, .~ status * fert * light)
m_comp_resource_yj <- update(m_yj, .~ status * comp_treat * light)
m_status <- update(m_yj, .~ status * status_comp)
```

```{r, eval = F}
u <- rep(1, nrow(dat_zz_only))
m_resource_zz <- update(m_zz, .~ status * fert * light)
m_comp_resource_zz <- update(m_zz, .~ status * comp_treat * light)
```


## origin:fert:light

```{r}
#function
plot_resource <- function(model = m_resource, ymax = 17, title = NA){
  t <- effect('status:fert:light', model) %>% as.data.frame() %>% 
    mutate(xaxis = c(1, 2, 4, 5, 7, 8, 10, 11))
  
  plot <- t %>% 
    mutate(origin = status) %>% # rename
    ggplot() +
    geom_point(aes(x = xaxis, y = fit %>% exp, col = origin),size = 2.5) +
    geom_errorbar(aes(ymin = lower %>% exp, ymax = upper %>% exp, x = xaxis, col = origin), 
                  width = 0,  size = 0.6,
                  inherit.aes = FALSE) +
    scale_x_continuous(limits = c(0.5, 11.5), breaks = c(1.5, 4.5, 7.5, 10.5), 
                       labels = c('Low resources', '+ nutrient', '+ light', '+ nutrient & light')) + 
    scale_y_continuous(limits = c(0, ymax)) + 
    scale_color_discrete(labels = c('non-native', 'native')) +
    labs(x = '', y = 'Aboveground biomass [g]\n', col = 'Origin') + theme
  
  if (!is.na(title)) {
    plot <- plot + labs(title = title)
  }
  return(plot)
}
```

### joint (Fig. 2)

```{r, fig.width= 12/2.54}
p2 <- plot_resource(m_resource) +
  annotate(geom = 'text', x = 1, y = 16, label = 'nutrient*', colour = 'black', hjust = 0, size = 2.4) +
  annotate(geom = 'text', x = 1, y = 14.75, label = 'light*', colour = 'black', hjust = 0, size = 2.4) +
  annotate(geom = 'text', x = 1, y = 13.5, label = 'origin x light x nutrient*', colour = 'black', hjust = 0, size = 2.4) +
  theme(legend.position = 'bottom')
p2

ggsave('./fig/f2_pooled.pdf', p2, height = 7.5/2.54, width = 9/2.54, useDingbats=FALSE)
```


### separate (Fig. s1)
```{r}

p_s1a <- plot_resource(m_resource_yj, ymax = 50, title = 'China\n' )   + labs(tag = 'a') + 
  theme(axis.text.x = element_text(size = 7),
        axis.text.y  = element_text(size = 7))
p_s1b <- plot_resource(m_resource_zz, ymax = 4,  title = 'Germany\n' ) + labs(tag = 'b') + 
  theme(axis.text.x = element_text(size = 7),
        axis.text.y  = element_text(size = 7))

ggsave('./fig/s1.pdf', grid.arrange(p_s1a, p_s1b, ncol =2, nrow = 1), 
       height = 8.5/2.54, width = 20/2.54, useDingbats=FALSE)

```

### different competition treatment (Fig. s4)
```{r, eval = F}
m_resource_comp <- update(m0, .~ status * fert * light*comp_treat, random = random)
```


```{r}

t_res_comp <- effect('status * fert * light*comp_treat', m_resource_comp) %>% as.data.frame() %>% 
  arrange(status, comp_treat) %>% 
  mutate(xaxis = rep(c(1:3, 5:7), each = 4))

```



```{r}
dat_p <- dat %>% 
  mutate(origin = status) %>% # rename
  mutate(cate = paste(light, fert),
         cate = ifelse(cate == 'low light low', 'Low resources', cate),
         cate = ifelse(cate == 'low light high', '+ nutrient', cate),
         cate = ifelse(cate == 'high light low', '+ light', cate),
         cate = ifelse(cate == 'high light high', '+ nutrient & light', cate)
         ) %>% 
  mutate(cate = factor(cate, levels = c('Low resources', '+ nutrient', '+ light', '+ nutrient & light'))) %>% 
  left_join(t_res_comp %>% dplyr::select(status, comp_treat, xaxis) %>% unique)
```

```{r}

col_fig <- c('#F39C12', '#115888', '#85C1E9')
t_res_comp %>% 
  mutate(origin = status) %>% # rename
  mutate(cate = paste(light, fert),
         cate = ifelse(cate == 'low light low', 'Low resources', cate),
         cate = ifelse(cate == 'low light high', '+ nutrient', cate),
         cate = ifelse(cate == 'high light low', '+ light', cate),
         cate = ifelse(cate == 'high light high', '+ nutrient & light', cate)
         ) %>% 
  mutate(cate = factor(cate, levels = c('Low resources', '+ nutrient', '+ light', '+ nutrient & light'))) %>% # order
  ggplot() +
  geom_jitter(data = dat_p, aes(x = xaxis, y = biomass),  colour = 'grey80', alpha = 0.4, width = 0.1, size = 1) + 
  geom_point(aes(x = xaxis, y = fit %>% exp, col = comp_treat),size = 2) +
  geom_errorbar(aes(ymin = lower %>% exp, ymax = upper %>% exp, x = xaxis, col = comp_treat), 
                width = 0,  size = 0.6,
                inherit.aes = FALSE) + 
  scale_x_continuous(breaks = c(2, 6), labels = c('Non-native', 'Native'), limits = c(0.5, 7.5)) +
  scale_color_manual(values = col_fig, labels = c('no', 'inter', 'intra')) + 
  labs(x = '', y = 'Aboveground biomass [g]\n', col = 'Competition') +
  theme + theme(strip.background = element_rect(fill = 'gray80', colour = 'black'),
                panel.grid = element_line(color = NA)) + 
  facet_wrap(~ cate)
ggsave('./fig/s4.pdf',
       height = 10/2.54, width = 15/2.54, useDingbats=FALSE)
```


## fert:light:competition (Fig. 3)
```{r}
t_comp <- effect('light * fert* comp_treat', m_comp) %>% as.data.frame() %>% 
  arrange(desc(light), desc(fert)) %>% 
  mutate(xaxis = c(1:3, 5:7, 9:11, 13:15))
```

```{r, fig.width= 12/2.54}
col_fig <- c('#F39C12', '#115888', '#85C1E9')

p3 <- t_comp %>% 
  ggplot() +
  geom_point(aes(x = xaxis, y = fit %>% exp, col = comp_treat),size = 2.5) +
  geom_errorbar(aes(ymin = lower %>% exp, ymax = upper %>% exp, x = xaxis, col = comp_treat), 
                width = 0,  size = 0.6,
                inherit.aes = FALSE) +
  scale_x_continuous(limits = c(0.5, 15.5), breaks = c(2, 6, 10, 14), 
                     labels = c('Low resources', '+ nutrient', '+ light', '+ nutrient & light')) + 
  scale_y_continuous(limits = c(0, 17)) + 
  scale_color_manual(values = col_fig,  labels = c('no', 'inter', 'intra')) + 
  labs(x = '', y = 'Aboveground biomass [g]\n', col = 'Competition') + theme +
  annotate(geom = 'text', x = 1, y = 16, label = 'competition x nutrient*', colour = 'black', hjust = 0, size = 2.4) +
  annotate(geom = 'text', x = 1, y = 14.75, label = 'competition x light*', colour = 'black', hjust = 0, size = 2.4) +
  annotate(geom = 'text', x = 1, y = 13.5, label = 'inter vs intra*', colour = 'black', hjust = 0, size = 2.4) +
  theme(legend.position = 'bottom')
p3
ggsave('./fig/f3_comp.pdf', p2, height = 7.5/2.54, width = 9/2.54, useDingbats=FALSE)
```



## origin:origin_comp (Fig. s2)
```{r}
t_status <- effect('status:status_comp', m_status) %>% as.data.frame() %>% 
  arrange(status)
t_status <- t_status[c(1,4, 5, 8),] %>% 
  mutate(xaxis = c(1:2, 4:5))

```

```{r, fig.width= 12/2.54}
p_comp <- t_status %>% 
  ggplot() +
  geom_point(aes(x = xaxis, y = fit %>% exp, col = status), 
             size = 2.5) +
  geom_errorbar(aes(ymin = lower %>% exp, ymax = upper %>% exp, x = xaxis, col = status), 
                width = 0,  inherit.aes = FALSE,
                size = 0.6) +
  scale_x_continuous(limits = c(0.5, 5.5), breaks = c(1, 2, 4, 5), 
                     labels = c('Non-native', 'Native', 'Non-native', 'Native')) + 
  scale_y_continuous(limits = c(0, 20)) + 
  labs(x = '\nOrigin of the competitor', y = 'Aboveground biomass [g]\n', 
       col = 'Origin of the target', title = 'China') + 
  scale_color_discrete(labels = c('non-native', 'native')) +
  theme +
  theme(legend.position = 'bottom')
p_comp

ggsave('./fig/s2_origin.pdf', p_comp, height = 8.5/2.54, width = 9/2.54, useDingbats=FALSE)
```

## origin:competition:light (Fig. S3)


```{r}
t_light_comp_zz <- effect('status:comp_treat:light', m_comp_resource_zz) %>% as.data.frame() %>% 
  filter(comp_treat != 'alone') %>% 
  arrange(comp_treat) %>% 
  mutate(xaxis = c(1, 2, 4, 5, 1, 2, 4, 5))
```


```{r}
p_light_comp_zz <- t_light_comp_zz %>% 
  mutate(comp_treat = ifelse(comp_treat == 'intra', 'Intraspecific competition', 'Interspecific competition')) %>% 
  ggplot() +
  geom_point(aes(x = xaxis, y = fit %>% exp, col = status), size = 2.5) +
  geom_errorbar(aes(ymin = lower %>% exp, ymax = upper %>% exp, x = xaxis, col = status), 
                width = 0,  size = 0.6,
                inherit.aes = FALSE) +
  scale_x_continuous(limits = c(0.5, 5.5), breaks = c(1.5, 4.5), 
                     labels = c('Low light-intensity', 'High light-intensity')) + 
  scale_y_continuous(limits = c(0, 4)) + facet_wrap( ~ comp_treat) + 
  labs(x = '', y = 'Aboveground biomass [g]\n', col = 'Origin', title = 'Germany') + 
  scale_color_discrete(labels = c('non-native', 'native')) +
  theme +
  theme(panel.spacing = unit(2, 'lines'),
        legend.position = 'bottom') 
p_light_comp_zz
ggsave('./fig/s3_origin_light_comp.pdf', p_light_comp_zz, height = 8.5/2.54, width = 12/2.54, useDingbats=FALSE)
```



```{r}
# save(file = 'results20221004.RData', list = ls())
```

